{ \sffamily \LARGE Electronic supplementary material}\\{\bf \sffamily \LARGE Evolution of reciprocity with limited payoff memory

### 2023-11-01 paper/si.tex
include expected payoff of one interaction in si

In the following, we discuss our different memory settings in more detail. 
We discuss five cases explicitly. In these cases, updating occurs
({\it i})~based on average payoffs based on all interactions (perfect memory), 
({\it ii})~based on the last round of one interaction (limited memory), 
({\it iii})~based on the last round of two interactions,
({\it iv})~based on the last two rounds of one interaction, and 
({\it v})~based on the last two rounds of two interactions. 
In each case, we consider the case that there are only two strategies present in the population (a resident and a mutant strategy). 
We first derive how likely it is that a learner (of any type) assigns a given payoff $\pi_\learner$ to itself, and a payoff of $\pi_\rolemodel$ to the role model. 
This allows us to derive explicit expressions for $\lambda^-_k/\lambda^+_k$, and hence for the mutant's fixation probability according to Eq.~\eqref{eq:fixation_probability}.
Based on these expressions we can characterize under which conditions cooperation is stochastically stable.

@@

~\\
{\bf Stochastic stability of cooperation.}
As an application of this formalism, we can compute when cooperation is stochastically stable in the perfect-information setting. 
To this end, suppose there is only a single mutant, $k\!=\!1$. The residents adopt Generous Tit-for-Tat, $\gtft = (1, 1, q)$ and the mutant adopts $\alld\! = \! (0,0,0)$. When two GTFT players interact, the resulting average distribution according to Eq.~\eqref{Eq:vAverage} simplifies to
\begin{align*}
\mathbf{v}(\gtft,\gtft) = (1, 0, 0, 0).
\end{align*}
On the other hand, if \alld{} interacts with \gtft, the respective probabilities become,
\begin{align*}
  \mathbf{v}(\alld,\gtft) = \big(0,~0,~1\!-\!\delta\!+\!\delta q,~\delta(1\! -\! q)\big).
\end{align*}
Based on Eq.~\eqref{Eq:ExpPay}, we can compute the strategies' expected payoffs as
\begin{align*}
\pi_\gtft =  \frac{N\!-\!2}{N-1} (b - c)  \!-\! \frac{1}{N-1}\big(1\!-\!\delta\!+\!\delta q\big)c
\quad \text{and} \quad
\pi_\alld  = \big(1\!-\!\delta\!+\!\delta q\big)b.
\end{align*}
As a consequence, we can calculate the corresponding ratio of transition probabilities according to Eq.~\eqref{eq:LambdaRatio},
\begin{align*}
\frac{\lambda_1^{-}}{\lambda_1^{+}} = e^{-\beta \left( (1-\delta+\delta q)(b+\frac{c}{N-1})- \frac{N-2}{N-1}(b-c) \right)}
\end{align*}
By definition, cooperation is stochastically stable if this ratio exceeds one, which is equivalent to
\begin{equation*}
q<1-\frac{1}{\delta}\cdot \frac{b+(N\!-\!1)c}{(N\!-\!1)b+c}.
\end{equation*}
For such a strategy to be feasible we require $q\!>\!0$, which implies $\delta \!>\! \big(b+(N\!-\!1)c\big)/\big((N\!-\!1)b\!+\!c\big)$.
In particular, in the limit of large populations \(N \!\rightarrow\! \infty\), we obtain that cooperation is stochastically stable if  \(q \!<\! 1\! -\! c/(\delta b)\). 
The minimum continuation probability for such a strategy to exist is $\delta \!>\! c/b$. 
In this way, we recover the classical conditions for cooperation to be feasible under direct reciprocity~\citep{molander:jcr:1985,Nowak1992tit,Schmid:NHB:2021}.

@@

\subsection{Updates based on the realized payoff one repeated game}\label{section:one_interaction_full_memory}

@@

{\bf Motivation.}
In the case of limited memory, we have thus far considered that individuals
update their strategies based solely on the number of interactions and by
considering only the payoff they received in those interactions in the final
turns. \\

\noindent
A player's payoff in such an interaction is calculated based on the
probabilities. In comparison, in the perfect memory case, individuals consider
their expected payoff from the interaction. A natural question to ask is what
happens if individuals update based on the expected payoffs of the interaction.

\noindent
{\bf Computing the ratio $\lambda^-_k/\lambda^+_k$.}
We consider the population setup once more, consisting of \(k\) mutants with the
strategy \(\strategy_\mutant=(y_\mutant,p_\mutant,q_\mutant)\), and \(N-k\)
residents with strategies \(\strategy_\resident =
(y_\resident,p_\resident,q_\mutant)\). In each step of the evolutionary process,
we select a learner and a role model. The learner or the role model can be
either a resident, and the other player is a mutant. As mentioned previously,
two major cases must be considered:

\begin{enumerate}
  \item The learner and the role model have had their most recent interaction
  with each other.
  \item The learner's last interaction was not with the role model. In the
  latter case, there are four different subcases based on whether the resident's
  last interaction partner was a mutant or a resident, and whether the mutant's
  last interaction partner was a mutant or a resident.
\end{enumerate}

\noindent
In each pairing, we assume that the players interact for a specific number of
rounds, and the payoff from their interaction is the average of the payoffs
described in Eq.~\eqref{eq:expected_payoff}. As a result, there are four
possible payoffs in our population setup: the payoffs an individual with a
resident strategy can receive when interacting with an individual with the same
strategy or with a mutant strategy, and similarly, the payoffs for the mutant
strategy. \\

@@

\begin{equation*}
  \setlength{\arraycolsep}{1pt} 
  \begin{array}{llll}
    \lambda^+	 &=
  &\frac{1}{N\!-\!1}\cdot  & \rho\left(\pi (\strategy_\mutant, \strategy_\resident), \pi(\strategy_\resident, \strategy_\mutant)\right)\\[0.5cm]
  &+
  &\frac{N\!-\!2}{N\!-\!1} \cdot 
  &\left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3}\cdot \rho(\pi(\strategy_\mutant, \strategy_\mutant), \pi(\strategy_\resident, \strategy_\mutant)) + 
   \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3}\cdot \rho(\pi(\strategy_\mutant, \strategy_\resident), \pi(\strategy_\resident, \strategy_\mutant)) \right.\\[0.5cm]
  &&&\left. +\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} \cdot \rho(\pi(\strategy_\mutant, \strategy_\mutant), \pi(\strategy_\resident, \strategy_\resident)) + 
   \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} \cdot \rho(\pi(\strategy_\mutant, \strategy_\resident), \pi(\strategy_\resident, \strategy_\resident),)\right].
  \end{array}
\end{equation*}

\noindent
{\bf Stochastic stability of cooperation.}
We once again calculate how easily a single \alld{} mutant can invade into a
resident population of \gtft{} players. When resident interacts with another
resident or with a mutant,
the expected payoffs based on Eq.~\eqref{eq:expected_payoff} of the interactions are,
\begin{equation*}
\begin{array}{ll}
  \pi (\strategy_\resident, \strategy_\resident) =  b - c,
  &\pi (\strategy_\resident, \strategy_\mutant) =  - c \left(\delta \left(q - 1\right) + 1\right).
\end{array}
\end{equation*}
The expected payoff of the mutant, are
\begin{align*}
\begin{array}{ll}
  \pi (\strategy_\mutant, \strategy_\resident) =  b \left(\delta \left(q - 1\right) + 1\right),
  &\pi (\strategy_\mutant, \strategy_\mutant) =  0.
\end{array}
\end{align*}
Based on these expected payoffs, we compute the ratio
\begin{equation}
  \frac{\lambda^{-}_1}{\lambda^{+}_1} =
  \frac{\frac{1}{N-1} + \frac{N-2}{N-1}}{\frac{1}{N-1} + \frac{N-2}{N-1}}
\end{equation}

@@

\begin{figure}[t!]
    \centering 
    \includegraphics[width=\textwidth]{static/more_memory_heatmaps_donation_game_with_illustrations.pdf}
    \caption{\textbf{Evolutionary dynamics of our five different settings.}
    From left to right, we present results on the following cases: individuals update their strategies based on
    ({\it i})~the expected payoffs (perfect memory); ({\it ii}) the last-round payoff
    of one interaction (limited memory); ({\it iii})~the last round payoff of two
    interactions; ({\it iv})~the payoffs of the last two rounds of one interaction; 
    and ({\it v}) the payoff of the last two rounds of two interactions. We run each simulation for \(T
    = 10^7\) time steps. For each time step we record the current resident
    population \((y, p, q)\). The graphs show how often the
    resident population chooses each combination \((p, q)\); because $\delta\!\approx\!1$, we do not show the distribution of the player's first-round cooperation probability $y$.}\label{fig:expected_payoffs_results}
    \end{figure}

{\bf Strategies favored by selection.} 
To gain some insights into the strategies evolving in each of the settings described in the previous section, we simulate the evolutionary process according to Algorithm~\ref{algorithm:pairwise_comparison}. 
Each time step, we record the current resident population \((y, p, q)\). 
The results of these simulations are shown in Supplementary Fig.~\ref{fig:expected_payoffs_results}. 
Qualitatively, all settings lead to similar dynamics as in the case of perfect and limited memory described in the main text. 
In particular, populations are clustered around \alld, and around a strip of conditionally cooperative strategies, \gtft$=\!(1,1,q)$. 
In all settings, we observe that along this strip, strategies are most abundant when they satisfy the condition for stochastic stability. 
For perfect memory, this means the resident population predominantly adopts a strategy with \(q\! \leq\! 1 \!-\!
c/b\!=\!0.9\) (for the parameters used in the simulation). 
For limited memory, the generosity of a conditional cooperator satisfies  \(q\!<\!\frac{1}{2}\). 
In the next two cases, when players either remember the last round of two interactions, or the last two rounds of one interaction, stochastic stability requires $q\!<\!\sqrt{2}{2}\!\approx\!0.701$. 
In all cases, the simulation results are well aligned with these theoretical predictions.\\

@@

\begin{figure}[!htbp]
	\centering 
	\includegraphics[width=.5\textwidth]{static/mutation_perfect_and_limited_memory_donation_game.pdf}
	\caption{\textbf{Evolutionary dynamics results for perfect and limited memory
			for different mutation values.}
		We perform five independent simulations. Simulations are run
		for $T\!=\!4\times 10^7$ time steps for each parameter. In each time step,
		we introduce a new mutant with a probability \(\mu\), and we then select
		two random players to serve as the role model and the learner. The learner
		adopts the strategy of the role model with a probability \(\rho(\pi_\learner, \pi_\rolemodel)\) where the
		payoffs depend on the relevant setting. We plot the average cooperation rate
		within the resident population for each value of \(\mu\). 
		The results show that our previous results for rare mutations hold more generally:
		While limited memory allows the evolution of moderate cooperation levels, perfect memory is always more conducive to the evolution of cooperation. 
		Parameters are the same as before, \(N =100, c=1, b=10, \beta=1\).}\label{fig:mutation}
\end{figure}

@@

\begin{figure}[!htbp]
  \centering 
  \includegraphics[width=\textwidth]{static/memory_one_results_low_benefit.pdf}
  \caption{\textbf{Evolution of memory-one strategies when the benefit of cooperation is comparably small.}
  We perform two independent simulations. In one simulation, individuals use
  expected payoffs. 
  In the other, they update their strategy based on the payoff they received in the last round of their last interaction. 
  We run each simulation for \(T = 10^8\) time steps.
  For each time step, we record the current resident population, which is
  now of the form  \((y, p_1, p_2, p_3, p_4)\). In the left panel, we report the
  cooperation rates for each simulation. 
  For memory-one strategies, like for reactive strategies, the use of expected payoffs results in more cooperation. 
  The right panel reports the most abundant strategy of each
  simulation. Abundance is the number of mutants a strategy can repel before
  being invaded. The most abundant strategies have some similarities,
  namely, \(p_1 \approx 1\), \(p_3 \approx 0\) and \(p_4 > \frac{1}{2}\). 
  Parameters: \(N =100, c=1, b=3, \beta=1\).}\label{fig:memory_one_low_benefit}
\end{figure}

@@

\begin{figure}[!htbp]
  \centering 
  \includegraphics[width=\textwidth]{static/cooperation_rate_over_b_and_beta_memory_one.pdf}
  \caption{{\bf Evolution of direct reciprocity for different parameter values.}}
\end{figure}

>>>>

In the following, we discuss our different memory settings in more detail. 
We discuss six cases explicitly. In these cases, updating occurs
({\it i})~based on average payoffs based on all interactions (perfect memory), 
({\it ii})~based on the last round of one interaction (limited memory), 
({\it iii})~based on the last round of two interactions,
({\it iv})~based on the last two rounds of one interaction,
({\it v})~based on the last two rounds of two interactions, and
({\it v})~based on the expected payoff of one interaction.
In each case, we consider the case that there are only two strategies present in the population (a resident and a mutant strategy). 
We first derive how likely it is that a learner (of any type) assigns a given payoff $\pi_\learner$ to itself, and a payoff of $\pi_\rolemodel$ to the role model. 
This allows us to derive explicit expressions for $\lambda^-_k/\lambda^+_k$, and hence for the mutant's fixation probability according to Eq.~\eqref{eq:fixation_probability}.
Based on these expressions we can characterize under which conditions cooperation is stochastically stable.

@@

~\\
{\bf Stochastic stability of cooperation.}
As an application of this formalism, we can compute when cooperation is stochastically stable in the perfect-information setting. 
To this end, suppose there is only a single mutant, $k\!=\!1$. The residents adopt Generous Tit-for-Tat, $\gtft = (1, 1, q)$ and the mutant adopts $\alld\! = \! (0,0,0)$. When two GTFT players interact, the resulting average distribution according to Eq.~\eqref{Eq:vAverage} simplifies to
\begin{align*}
\mathbf{v}(\gtft,\gtft) = (1, 0, 0, 0).
\end{align*}
On the other hand, if \alld{} interacts with \gtft, the respective probabilities become,
\begin{align*}
  \mathbf{v}(\alld,\gtft) = \big(0,~0,~1\!-\!\delta\!+\!\delta q,~\delta(1\! -\! q)\big).
\end{align*}
Based on Eq.~\eqref{Eq:ExpPay}, we can compute the strategies' expected payoffs as
\begin{align*}
\pi_\gtft =  \frac{N\!-\!2}{N-1} (b - c)  \!-\! \frac{1}{N-1}\big(1\!-\!\delta\!+\!\delta q\big)c
\quad \text{and} \quad
\pi_\alld  = \big(1\!-\!\delta\!+\!\delta q\big)b.
\end{align*}
As a consequence, we can calculate the corresponding ratio of transition probabilities according to Eq.~\eqref{eq:LambdaRatio},
\begin{align*}
\frac{\lambda_1^{-}}{\lambda_1^{+}} = e^{-\beta \left( (1-\delta+\delta q)(b+\frac{c}{N-1})- \frac{N-2}{N-1}(b-c) \right)}
\end{align*}
By definition, cooperation is stochastically stable if this ratio exceeds one, which is equivalent to
\begin{equation}\label{Eq:Condition_PerfectMemory}
q<1-\frac{1}{\delta}\cdot \frac{b+(N\!-\!1)c}{(N\!-\!1)b+c}.
\end{equation}
For such a strategy to be feasible we require $q\!>\!0$, which implies $\delta \!>\! \big(b+(N\!-\!1)c\big)/\big((N\!-\!1)b\!+\!c\big)$.
In particular, in the limit of large populations \(N \!\rightarrow\! \infty\), we obtain that cooperation is stochastically stable if  \(q \!<\! 1\! -\! c/(\delta b)\). 
The minimum continuation probability for such a strategy to exist is $\delta \!>\! c/b$. 
In this way, we recover the classical conditions for cooperation to be feasible under direct reciprocity~\citep{molander:jcr:1985,Nowak1992tit,Schmid:NHB:2021}.

@@

\subsection{Updates based on expected payoff of one repeated game}\label{section:one_interaction_full_memory}

@@

{\bf Motivation.}

\noindent
{\bf Computing the ratio $\lambda^-_k/\lambda^+_k$.}
We revisit the population setup, consisting of \(k\) mutants with the strategy
\(\strategy_\mutant = (y_\mutant, p_\mutant, q_\mutant)\) and \(N-k\) residents
with strategies \(\strategy_\resident = (y_\resident, p_\resident,
q_\resident)\). In each step of the evolutionary process, we select a learner and
a role model. The learner or the role model can be either a resident, and the
other player is a mutant. We consider that the learner and the
model update based on their most recent interaction. As mentioned previously,
two major cases must be considered:

\begin{enumerate}
  \item The learner and the role model have had their most recent interaction with each other.
  \item The learner's last interaction was not with the role model.
\end{enumerate}

In the latter case, there are four different subcases based on whether the
resident's last interaction partner was a mutant or a resident and whether the
mutant's last interaction partner was a mutant or a resident.

We assume that the players' payoff received from their last interaction is their
expected payoff given by Eq.~\eqref{eq:expected_payoff}. Thus, a resident
interacting with a resident receives the payoff $\pi(s_\resident, s_\resident)$,
and $\pi(s_\resident, s_\mutant)$ when interacting with a mutant. Similarly, a
mutant interacting with a resident receives the payoff $\pi(s_\mutant,
s_\resident)$ and $\pi(s_\mutant, s_\mutant)$ when interacting with a mutant.\\

@@

\begin{equation*}
  \setlength{\arraycolsep}{1pt} 
  \begin{array}{llll}
    \lambda^-	 &=
  &\frac{1}{N\!-\!1}\cdot  & \rho\left(\pi (\strategy_\mutant, \strategy_\resident), \pi(\strategy_\resident, \strategy_\mutant)\right)\\[0.5cm]
  &+
  &\frac{N\!-\!2}{N\!-\!1} \cdot 
  &\left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3}\cdot \rho(\pi(\strategy_\mutant, \strategy_\mutant), \pi(\strategy_\resident, \strategy_\mutant)) + 
   \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3}\cdot \rho(\pi(\strategy_\mutant, \strategy_\resident), \pi(\strategy_\resident, \strategy_\mutant)) \right.\\[0.5cm]
  &&&\left. +\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} \cdot \rho(\pi(\strategy_\mutant, \strategy_\mutant), \pi(\strategy_\resident, \strategy_\resident)) + 
   \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} \cdot \rho(\pi(\strategy_\mutant, \strategy_\resident), \pi(\strategy_\resident, \strategy_\resident),)\right].
  \end{array}
\end{equation*}

\noindent
{\bf Stochastic stability of cooperation.}
We once again calculate how easily a single \alld{} mutant can invade into a
resident population of \gtft{} players. When resident interacts with another
resident or with a mutant,
the expected payoffs based on Eq.~\eqref{eq:expected_payoff} of the interactions are,
\begin{equation*}
\begin{array}{ll}
  \pi (\strategy_\resident, \strategy_\resident) =  b - c,
  &\pi (\strategy_\resident, \strategy_\mutant) =  - c \left(\delta \left(q - 1\right) + 1\right).
\end{array}
\end{equation*}
The expected payoff of the mutant, are
\begin{align*}
\begin{array}{ll}
  \pi (\strategy_\mutant, \strategy_\resident) =  b \left(\delta \left(q - 1\right) + 1\right),
  &\pi (\strategy_\mutant, \strategy_\mutant) =  0.
\end{array}
\end{align*}
Based on these expected payoffs, we compute the ratio
\begin{equation}
  \frac{\lambda^{-}_1}{\lambda^{+}_1} =
  \frac{\frac{1}{N-1}\frac{1}{\left(1 + e^{\beta \left(b \left(\delta q - \delta\right) + c\right)}\right)}
  + \frac{N - 2}{N-1}\frac{1}{\left(1 + e^{- \beta \left(- b \left(\delta q - \delta + 1\right) + b - c\right)}\right)}}
  {\frac{1}{N-1}\frac{1}{\left(1 + e^{- \beta \left(b \left(\delta q - \delta\right) + c\right)}\right)}
  + \frac{N - 2}{N-1}\frac{1}{\left(1 + e^{- \beta \left(b \left(\delta q - \delta + 1\right) - b + c\right)}\right)}}.
\end{equation}

\noindent
In the limit of large populations \(N \rightarrow \infty \), this ratio simplifies to

@@

\begin{figure}[t!]
    \centering 
    \includegraphics[width=\textwidth]{static/more_memory_heatmaps_donation_game_with_illustrations.pdf}
    \caption{\textbf{Evolutionary dynamics of our six different settings.}
    From left to right, we present results on the following cases: individuals update their strategies based on
    ({\it i})~the expected payoffs (perfect memory); ({\it ii}) the last-round payoff
    of one interaction (limited memory); ({\it iii})~the last round payoff of two
    interactions; ({\it iv})~the payoffs of the last two rounds of one interaction; 
    ({\it v}) the payoff of the last two rounds of two interactions and
    ({\it vi}) the expected payoff of one interaction.
    Simulations are run either for a comparably high benefit of cooperation~($b/c\!=\!10$), or for a low benefit~($b/c\!=\!3$).
    We run each simulation for \(T= 10^7\) time steps. For each time step we record the current resident
    population \((y, p, q)\). The graphs show how often the
    resident population chooses each combination \((p, q)\); because $\delta\!\approx\!1$, we do not show the distribution of the player's first-round cooperation probability $y$.}\label{fig:expected_payoffs_results}
    \end{figure}

{\bf Strategies favored by selection.} 
To gain some insights into the strategies evolving in each of the settings described in the previous section, we simulate the evolutionary process according to Algorithm~\ref{algorithm:pairwise_comparison}. 
Each time step, we record the current resident population \((y, p, q)\). 
The results of these simulations are shown in Supplementary Fig.~\ref{fig:expected_payoffs_results}. 
Qualitatively, all settings lead to similar dynamics as in the case of perfect and limited memory described in the main text. 
In particular, populations are clustered around \alld, and around a strip of conditionally cooperative strategies, \gtft$=\!(1,1,q)$. 
In all settings, we observe that along this strip, strategies are most abundant when they satisfy the condition for stochastic stability. 
For perfect memory, this means the resident population predominantly adopts a
strategy with \(q \leq 1 - \frac{c}{b} = 0.9\) for high benefit and \(q \leq 1 -
\frac{c}{b} = 0.67\) for low benefit.
For limited memory, the generosity of a conditional cooperator satisfies \(q <
\frac{1}{2}\) for both values of benefit.
In the next two cases, when players either remember the last round of two
interactions or the last two rounds of one interaction, stochastic stability
requires \(q < \frac{\sqrt{2}}{2} \approx 0.701\).
In the final case, we observe that the generosity of a conditional cooperator is
in line with the condition for stochastic stability in the perfect memory case:
\(q < 1 - \frac{c}{b} = 0.9\) for high benefit and \(q < 1 - \frac{c}{b} =
0.67\) for low benefit.
In all cases, the simulation results are well aligned with these theoretical predictions.\\

@@

\begin{figure}[!htbp]
	\centering 
	\includegraphics[width=.5\textwidth]{static/mutation_perfect_and_limited_memory_donation_game.pdf}
	\caption{\textbf{Evolutionary dynamics results for perfect and limited memory
			for different mutation values.}
		We perform five independent simulations. Simulations are run
		for $T\!=\!4\times 10^7$ time steps for each parameter. In each time step,
		we introduce a new mutant with a probability \(\mu\), and we then select
		two random players to serve as the role model and the learner. The learner
		adopts the strategy of the role model with a probability \(\rho(\pi_\learner, \pi_\rolemodel)\) where the
		payoffs depend on the relevant setting. We plot the average cooperation rate
		within the resident population for each value of \(\mu\). 
		The results show that our previous results for rare mutations hold more generally:
		While limited memory allows the evolution of moderate cooperation levels, perfect memory is always more conducive to the evolution of cooperation. 
		Parameters are the same as before, \(N =100, c=1, b=10, \beta=1, \delta=0.999\).}\label{fig:mutation}
\end{figure}

@@

To further investigate the influence of different parameters, we have
systematically varied the benefit~$b$ and the selection strength~$\beta$, as
shown in Fig.~\ref{fig:cooperation_rate_over_benefit_and_beta_memory_one}.
When compared to the case of reactive strategies, what we observe
here is that in the case of limited payoff memory, cooperation rates increase
with an increasing benefit~$b$ and do not remain stable
(Fig.~\ref{fig:cooperation_rate_over_benefit_and_beta_memory_one}(a)). However,
even in this scenario, we still observe that perfect memory consistently results
in a higher cooperation rate.
Regarding the effect of different selection strengths, as depicted in
Fig.~\ref{fig:cooperation_rate_over_benefit_and_beta_memory_one}(b), the results
once again suggest that both perfect and limited payoff memory yield similar
cooperation rates for weak selection (\(\beta \!<\! 1\)). Beyond weak selection, increasing
selection has a positive effect under perfect payoff memory but a negative
effect under limited payoff memory.

\begin{figure}[!htbp]
  \centering 
  \includegraphics[width=\textwidth]{static/memory_one_results_low_benefit.pdf}
  \caption{\textbf{Evolution of memory-one strategies when the benefit of cooperation is comparably small.}
  We perform two independent simulations. In one simulation, individuals use
  expected payoffs. 
  In the other, they update their strategy based on the payoff they received in the last round of their last interaction. 
  We run each simulation for \(T = 10^8\) time steps.
  For each time step, we record the current resident population, which is
  now of the form  \((y, p_1, p_2, p_3, p_4)\). In the left panel, we report the
  cooperation rates for each simulation. 
  For memory-one strategies, like for reactive strategies, the use of expected payoffs results in more cooperation. 
  The right panel reports the most abundant strategy of each
  simulation. Abundance is the number of mutants a strategy can repel before
  being invaded. The most abundant strategies have some similarities,
  namely, \(p_1 \approx 1\), \(p_3 \approx 0\) and \(p_4 > \frac{1}{2}\). 
  Parameters: \(N =100, c=1, b=3, \beta=1, \delta=0.999\).}\label{fig:memory_one_low_benefit}
\end{figure}

@@

\begin{figure}[!htbp]
  \centering 
  \includegraphics[width=\textwidth]{static/cooperation_rate_over_b_and_beta_memory_one.pdf}
  \caption{{\bf Evolution of direct reciprocity for different parameter values for memory-one strategies.}
  To further explore the robustness of our results in the case of memory-one
  strategies, we run simulations for different benefit values (left panels, A)
  and for different selection strengths (right panels, B). These simulations
  suggest that perfect payoff memory consistently leads to more cooperation and
  more generosity. Unless explicitly varied, the parameters of the simulation
  are $N\!=\!100$, $b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.999$.
  Simulations are run for $T\!=\!5\times 10^7$ time steps for each parameter
  combination.}
  \label{fig:cooperation_rate_over_benefit_and_beta_memory_one}
\end{figure}
